---
title: "p8105_hw2_yl6109"
author: "Yurou Liu"
date: "2025-09-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### pols-month
```{r}
library(dplyr)
library(readr)
library(tidyr)


pols_df = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names()


pols_df_sep = pols_df %>% 
  separate(mon, c("year", "month", "day")) 
pos_df_month = pols_df_sep %>% 
  mutate(month = recode(month, `01` = "January", `02` = "February", `03` = "March", `04` = "April", `05` = "May", `06` = "June", `07` = "July", `08` = "August", `09` = "September", `10` = "October", `11` = "November", `12` = "December"))


month_pols = pos_df_month %>% 
  mutate(president = ifelse(prez_gop > prez_dem, "gop", "dem")) 


month2_pols = select(month_pols, -prez_dem, -prez_gop, -day)
```


### snp

```{r}
snp_df = read_csv("./data/snp.csv") %>% 
  janitor::clean_names()


sep_snp = snp_df %>%
  separate(date, c("date", "month", "year"))


sep2_snp = sep_snp %>% 
  arrange(month)


month_snp = sep2_snp %>% 
  mutate(month = recode(month, `1` = "January", `2` = "February", `3` = "March", `4` = "April")) 


month_snp$year = as.numeric(sep_snp$year)


month2_snp = month_snp %>% 
  mutate(year = ifelse(year <= 25, 2000 + year, 1900 + year)) %>% 
  arrange(year) %>% 
  select(year, month, everything()) %>% 
  select(-date)
```

### unemployment

```{r}
unemployment_df = read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec, 
    names_to = "month",
    values_to = "unemployment_rate(%)"
  )


month_un = unemployment_df %>% 
  mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July", "aug" = "August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" = "December"))
```


### merging the three datasets

```{r}
month2_pols$year = as.numeric(month2_pols$year)

fianl_df = month2_pols %>% 
  left_join(month2_snp, by = c("year", "month")) %>% 
  left_join(month_un, by = c("year", "month"))
```

### description


## Problem 2

```{r}
library(readxl)

mr_tw_df = read_excel("./data/wheel.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(-x15, -x16) %>%
  filter(row_number() != 708) %>% 
  mutate(trash_wheel_type = "Mr")


prof_tw_df = read_excel("./data/wheel.xlsx", sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>% 
  filter(row_number() != 133) %>% 
  mutate(trash_wheel_type = "Professor")


gf_tw_df = read_excel("./data/wheel.xlsx", sheet = "Gwynns Falls Trash Wheel") %>% 
  janitor::clean_names() %>% 
  filter(row_number() != 350) %>% 
  mutate(trash_wheel_type = "Gwynns_Falls")


mr_tw_df$year = as.numeric(mr_tw_df$year)
final_df = bind_rows(mr_tw_df, prof_tw_df, gf_tw_df) %>% 
  select(year, month, date, trash_wheel_type, dumpster, everything()) %>%
  arrange(year, month, date)
```

### Description

## Problem 3

construct the merged dataframe.

```{r}
zc_df = read_csv("./data/Zip Codes.csv") %>% 
  janitor::clean_names() %>% 
  select(-file_date, -county_code, -state_fips)
  

znyc_df = read_csv("./data/Zip_NYC.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    x2015_01_31:x2024_08_31,
    names_to = "date",
    names_prefix = "x",
    values_to = "price") %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(county_name = recode(county_name, "Bronx County" = "Bronx", "Kings County" = "Kings", "New York County" = "New York", "Queens County" = "Queens", "Richmond County" = "Richmond")) %>%
  mutate(county = county_name, zip_code = region_name) %>% 
  select(-county_name, -region_name, -state_name) 


merged_df = zc_df %>%
  left_join(znyc_df, by = "zip_code", "county") %>% 
  mutate(county = county.x) %>%
  select(-county.x, -county.y, -region_type) %>% 
  arrange(year, month, day) %>% 
  mutate(month = recode(month, `01` = "January", `02` = "February", `03` = "March", `04` = "April", `05` = "May", `06` = "June", `07` = "July", `08` = "August", `09` = "September", `10` = "October", `11` = "November", `12` = "December")) %>% select(year, month, day, state, city, county, county_fips, neighborhood, region_id, zip_code, size_rank, price, everything()) 

final_merged_df = merged_df %>% 
  filter(row_number() < 17517)
```


count number of unique observations

```{r}
library(dplyr)

# for different zip codes observed,
unique_zc_count = 
  final_merged_df %>% 
  summarize(n_distinct(zip_code))

# for different neighborhoods observed,
unique_nbhd_count = 
  final_merged_df %>% 
  summarize(n_distinct(neighborhood))

# for different exact zip codes that are not included in zillow datasets,
zc_not_included = merged_df %>% 
  filter(row_number() >= 17517 & row_number() <= 17687) %>% 
  distinct(zip_code) %>%
  arrange(zip_code)

unique_zc_count %>% print()
unique_nbhd_count %>% print()
zc_not_included %>% print()
```
in total, 17516 observations exists. 149 unique zip codes are included. 43 unique neighborhoods are included.

according to merged_df with na year, zip codes that are not in the Zillow dataset can be found using the r code above.

For price difference between Jan 2020 and Jan 2021,
```{r}
library(tidyr)

y20_data = final_merged_df %>% 
  filter(month == "January" & year == "2020") %>% 
  select(zip_code, price, neighborhood) %>% 
  rename(y20_price = price)

y21_data = final_merged_df %>% 
  filter(month == "January" & year == "2021") %>% 
  select(zip_code, price, neighborhood) %>%
  rename(y21_price = price)

comparison = y20_data %>% 
  inner_join(y21_data, by = "zip_code") %>% 
  filter(!is.na(y20_price) & !is.na(y21_price)) %>% 
  mutate(price_dif = y20_price - y21_price) %>% 
  mutate(neighborhood = neighborhood.x) %>% 
  select(-neighborhood.x, -neighborhood.y)

largest_drop = comparison %>% 
  arrange(price_dif) %>% 
  head(10) %>% 
  select(zip_code, neighborhood, y20_price, y21_price, price_dif)

# the zip codes and corresponding neighborhoods of top 10 largest drops in rent price from 2020 to 2021 is shown below.
largest_drop %>% print()
```